---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
```

# tbat

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of `tbat` is to download data from The Banks Association of Turkey webpage. The webpage has a Data Query System but has no public API. `tbat` uses POST requests to download data programmatically.

## Installation

You can install the development version of tbat from [Github](https://github.com/emraher/tbat) with:

``` r
devtool::install_github("emraher/tbat")
```

## Usage

### Regional Data

There are 3 functions in this section.

- `years_region()`: Requires no inputs. Returns available years.

- `params_region()`: Requires no inputs. Returns parameter and region codes.

- `data_region(years = NULL, regions = NULL, parameters = NULL)`: Inputs are optional. Returns regional data.


```{r}
library(tbat)
        
(df <- params_region())

summary(years_region())
```

If no input argument is given, the function downloads all regional data. It may take some time to download all data. 

You can define `years`, `regions`, and/or `parameters`. To check available years, regions, and parameters first run `years_region()` and/or `params_region()`.

- `regions` are codes from `il_bolge_key` column of `params_region()` result.

- `parameters` are codes from `parametre_uk` column of `params_region()` result.



```{r}
(df <- data_region(years = 2018:2019, 
                   regions = c("-363956677", "1965766610"), 
                   parameters = c("4978", "11978")))
```


### Historical Data

There are 2 functions in this section.

- `bank_hist()`: Requires no inputs. Returns historical information about banks.

- `gm_hist()`: Requires no inputs. Returns historical information about managers.

```{r}
(bankHist <- bank_hist())

(gmHist <- gm_hist())
```


### Current Infomation Data

There are 2 functions in this section.

- `banks_summary()`: Requires no inputs. Returns number of currently operating banks and branches.

- `banks_info(date = format(Sys.time(), "%d/%m/%Y"))`: `date` is optional. If date (should be in DD/MM/YYYY format) is given returns info for the given date. Returns contact and information about banks.

```{r}
(bankSum <- banks_summary())

(bankInfo <- banks_info())
```

### Risk Center Data

There are 3 functions in this section.

- `periods_risk()`: Requires no inputs. Returns available periods.

- `categories_risk()`: Requires no inputs. Returns parameter codes.

- `data_risk(categories = NULL, periods = NULL)`: Inputs are optional. Returns risk data.

To check available periods and categories first run `periods_risk()` and/or `categories_risk()`.

- `periods` are codes from `DONEM` column of `periods_risk()` result.

- `categories` are codes from `uk_kategori` column of `categories_risk()` result.


```{r}
(periodsRisk <- periods_risk())

(categoriesRisk <- categories_risk())

(dataRisk <- data_risk(categories = c(-285334560, 2008209749, 949943133),
                       periods = c("2014-08", "2019-12")))

# -------------------------------------------------------------------------- ###
# Another Example using data from categoriesRisk and periodsRisk
# -------------------------------------------------------------------------- ###

mycategories <- categoriesRisk %>% 
  dplyr::filter(uk_rapor == 621177206) %>% 
  dplyr::select(uk_kategori) %>% 
  unlist()

myperiods <- periodsRisk %>% 
  dplyr::slice_sample(n = 3) %>% 
  unlist()

(dataRisk <- data_risk(categories = mycategories,
                      periods = myperiods))
```


### Financial Tables Data

There are 4 functions in this section.

- `periods_financial()`: Requires no inputs. Returns available periods.

- `banks_financial()`: Requires no inputs. Returns bank codes.

- `groups_financial()`: Requires no inputs. Returns bank group codes.

- `tables_financial(bank_code, periods = NULL)`: Returns table codes.

- `data_financial(periods, tables, banks = NULL, groups = NULL)`: Returns financial tables data. At least a bank or bank group must be chosen. 

    -  `periods`: Period codes from `periods_financial()` **key** column
    
    -  `tables`: Table codes from `tables_financial()` **unique_key** column
    
    -  `banks`: Bank codes from `banks_financial()` **banka_kodu** column
    
    -  `groups`: Group codes from `groups_financial()` **grup_no** column

```{r}
(periodsFinancial <- periods_financial())

(banksFinancial <- banks_financial())

(groupsFinancial <- groups_financial())

(tablesFinancial <- tables_financial(bank_code = c(3, 5)))

(dataFinancial <- data_financial(periods = c(1477300252),
                                 tables = c(2086),
                                 banks = c(3)))


(dataFinancial2 <- data_financial(periods = c(1477300252),
                                  tables = c(2087),
                                  groups = c(1)))
```


```{r, eval=FALSE}
(dataFinancial3 <- data_financial(periods = c(1477300252),
                                  tables = c(2087),
                                  banks = c(3),
                                  groups = c(1)))

#> Error in data_financial(periods = c(1477300252), tables = c(2087), banks = c(3),  : 
#> Cannot select both banks and groups at the same time. I'm working on this and will fix it soon!

```

